Modal type theory based on the intuitionistic epistemic logic

### 2018-01-24 article.tex
small updates

  At first we define the natural deduction for IEL$^{-}$ with $\textbf{K}$-modality and binary connectives
  $\to$ and $\land$ (we call that calculus NIEL$^{-}_{\land, \to}$):

  \begin{defin} Natural deduction NIEL$^{-}_{\land, \to}$ for IEL$^{-}$ with $\to$ and $\wedge$:

  \begin{center}
  \begin{prooftree}
  \AxiomC{}
  \RightLabel{\scriptsize{ax}}
  \UnaryInfC{$\Gamma , A \vdash A$}
  \end{prooftree}
  \end{center}

  \begin{minipage}{0.5\textwidth}
    \begin{flushleft}
  	\begin{prooftree}
  	\AxiomC{$\Gamma, A \vdash B$}
  	\RightLabel{$\rightarrow_i$}
  	\UnaryInfC{$\Gamma \vdash A \to B$}
    	\end{prooftree}

  	\begin{prooftree}
  	\AxiomC{ $\Gamma \vdash A$ }
  	\AxiomC{ $\Gamma \vdash B$ }
  	\RightLabel{$\land_i$}
  	\BinaryInfC{$\Gamma \vdash A \land B$}
  	\end{prooftree}

  	\begin{prooftree}
    	\AxiomC{$\Gamma \vdash A$}
    	\RightLabel{$\textbf{K}_I$}
    	\UnaryInfC{$\Gamma \vdash \textbf{K}A$}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
    	\begin{prooftree}
  	\AxiomC{$\Gamma \vdash A \to B$}
  	\AxiomC{$\Gamma \vdash A$}
  	\RightLabel{$\rightarrow_i$}
  	\BinaryInfC{$\Gamma \vdash B$}
    	\end{prooftree}

  	\begin{prooftree}
  	\AxiomC{ $\Gamma \vdash A_1 \land A_2$ }
  	\RightLabel{$\land_e, i \in \{ 1, 2 \}$}
  	\UnaryInfC{$\Gamma \vdash A_i$}
  	\end{prooftree}

  	\begin{prooftree}
    \AxiomC{$\Gamma \vdash {\bf K} \vec{A}$}
    \AxiomC{$\vec{A} \vdash B$}
    \BinaryInfC{$\Gamma \vdash {\bf K} B$}
  	\end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

@@

It is clearly that these lemmas could be extended for IEL$^{-}$ with $\vee$ and $\neg$ similary.

@@

      \begin{prooftree}
        \AxiomC{$\Gamma \vdash x : A$}
        \RightLabel{$\textbf{K}_I$}
        \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: x : {\bf K}A $}
      \end{prooftree}
  \end{minipage}%
  \hfill
  \begin{minipage}{0.45\textwidth}
  \begin{tabular}{p{\textwidth}}
    \begin{prooftree}
    \AxiomC{$\Gamma \vdash f : A \to B$}
    \AxiomC{$\Gamma \vdash x : A$}
    \RightLabel{$\rightarrow_e$}
    \BinaryInfC{$\Gamma \vdash fx : B$}
    \end{prooftree}

@@

  $\text{let}_{{\bf K}}$ is the same as the $\Box$-rule in typed lambda calculus for intuitionistic normal modal logic {\bf IK}, which is described in \cite{ModalK}.

@@

Here are some examples of derivable terms:

1) $\vdash (\lambda x. {\bf pure \: } x) : A \to \textbf{K}A$;

2) $\vdash \lambda f. \lambda x. {\bf let \: pure} \: \langle g, y \rangle = \langle f, x \rangle {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$

3) $\lambda f. \lambda x. {\bf let \:pure} \: \langle g, y \rangle = \langle {\bf pure \:} f, x \rangle {\: \bf in \: } g y : (A \to B) \to {\bf K} A \to {\bf K} B$

@@

For cases with application, abstraction and pairs see \cite{Morten} \cite{Pierce}.

@@

By generation $\Gamma \vdash M : A$ and $\Gamma \vdash N : A$ by assumption.
\\ So $\Gamma \vdash {\bf pure \:} N : {\bf K}A$.

\section{Strong normalization}

\section{Confluence}

@@

  A lax monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations:

@@

  2) $\ast_{A, B} : \mathcal{F}A \otimes_2 \mathcal{F}B \to \mathcal{F}(A \otimes_1 B)$

@@

  An applicative functor is a triple $\langle \mathcal{C}, \mathcal{K}, \eta \rangle$,
where $\mathcal{C}$ is a symmetric monoidal category, $\mathcal{K}$ is a lax monoidal endofunctor and $\eta$ is a natural transformation, such that:

@@

2) $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$.
\end{defin}

@@

\subsection{Soundness}

@@

\subsection{Completeness}

@@

\begin{theorem} Completeness
  $ $

  i) {\bf K} is an applicative functor;

  ii) Let $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$, then $M =_{\beta \eta} N$.
\end{theorem}

>>>>

  At first we define the natural deduction for IEL$^{-}$ :

  \begin{defin} Natural deduction NIEL for IEL$^{-}$ is an extenstion of intuitionistic natural deduction with additional derivation rules for modality:

  \begin{minipage}{0.5\textwidth}
    \begin{flushleft}
  	\begin{prooftree}
    	\AxiomC{$\Gamma \vdash A$}
    	\RightLabel{$\textbf{K}_I$}
    	\UnaryInfC{$\Gamma \vdash \textbf{K}A$}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
  	\begin{prooftree}
    \AxiomC{$\Gamma \vdash {\bf K} \vec{A}$}
    \AxiomC{$\vec{A} \vdash B$}
    \BinaryInfC{$\Gamma \vdash {\bf K} B$}
  	\end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

@@

@@

      \begin{prooftree}
        \AxiomC{$\Gamma \vdash M : A$}
        \RightLabel{$\textbf{K}_I$}
        \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: M : {\bf K}A $}
      \end{prooftree}
  \end{minipage}%
  \hfill
  \begin{minipage}{0.45\textwidth}
  \begin{tabular}{p{\textwidth}}
    \begin{prooftree}
    \AxiomC{$\Gamma \vdash M : A \to B$}
    \AxiomC{$\Gamma \vdash N : A$}
    \RightLabel{$\rightarrow_e$}
    \BinaryInfC{$\Gamma \vdash MN : B$}
    \end{prooftree}

@@

  $\text{let}_{{\bf K}}$ is similar to the $\Box$-rule in typed lambda calculus for intuitionistic normal modal logic {\bf IK}, which is described in \cite{ModalK}.

@@

Here are some examples of closed terms:

\begin{itemize}
\item $(\lambda x. {\bf pure \: } x) : A \to \textbf{K}A$;
\item $\lambda f. \lambda x. {\bf let \: pure} \: \langle g, y \rangle = \langle f, x \rangle {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$
\item $\lambda f. \lambda x. {\bf let \:pure} \: \langle g, y \rangle = \langle {\bf pure \:} f, x \rangle {\: \bf in \: } g y : (A \to B) \to {\bf K} A \to {\bf K} B$
\end{itemize}

@@

\begin{proof}
For cases with application, abstraction and pairs see \cite{Morten} \cite{Pierce}.

@@

By generation $\Gamma \vdash M : A$ and $\Gamma \vdash N : A$ by assumption.
\\ So $\Gamma \vdash {\bf pure \:} N : {\bf K}A$.
\end{proof}

Strong normalization and colfuence for ${\bf IK}$ was proved by Kakutani for call-by-value and for call-by name \cite{ModalK} \cite{ModalK1}.

@@

  A monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations, which satisfy the well-known conditions described in \cite{ElKelly}:

@@

  2) $\ast_{A, B} : \mathcal{F}A \otimes_2 \mathcal{F}B \to \mathcal{F}(A \otimes_1 B)$.

@@

  An applicative functor is a triple $\langle \mathcal{C}, \mathcal{K}, \eta \rangle$,
where $\mathcal{C}$ is a symmetric monoidal category, $\mathcal{K}$ is a monoidal and $\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{K}$ is a natural transformation (similar to unit in monad), such that:

@@

2) $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$;

3) Weak commutativity condition:

\xymatrix{
&& A \otimes \mathcal{K}B \ar[d]_{\sigma_{A, \mathcal{K}B}} \ar[rr]^{\eta_{A} \otimes id_{\mathcal{K}B}} && \mathcal{K}A \otimes \mathcal{K}B \ar[rr]^{\ast_{A,B}} && \mathcal{K}(A \otimes B) \ar[d]^{\mathcal{K}(\sigma_{A,B})}\\
&& \mathcal{K}B \otimes A \ar[rr]_{id_{\mathcal{K}B} \otimes \eta_{A}} && \mathcal{K}B \otimes \mathcal{K}A \ar[rr]_{\ast_{B,A}} && \mathcal{K}(B \otimes A)
}
\end{defin}

@@

\subsection{Soundness and completeness}

@@

\begin{theorem} Completeness

Let $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$, then $M =_{\beta \eta} N$.
\end{theorem}

@@

\begin{lemma}

  {\bf K} is an applicative functor
\end{lemma}
